\setstretch{1.05} SUPPORTING INFORMATION:\\ Towards atomistic resolution structure of phosphatidylcholine headgroup and glycerol backbone at different ambient conditions\footnote{Publication about results presented in the NMRlipids project (\url{nmrlipids.blogspot.fi})}

### 2015-11-23 HGmodelMANUSCRIPT/HGmodel_ACSsupplementary.tex
Compile pdf from changes in commits https://github.com/NMRLipids/nmrlipids.blogspot.fi/commit/a3bc5f847a9a6818302cd8919b06ba6aab8d7542 https://github.com/NMRLipids/nmrlipids.blogspot.fi/commit/c79f4092bfa48f8b9c046fa7a507b513e4b0ff12 https://github.com/NMRLipids/nmrlipids.blogspot.fi/commit/749c796304a5858b555b184e6e67cdfd9b763238 https://github.com/NMRLipids/nmrlipids.blogspot.fi/commit/467d3870a2569038ebf7dfef8e29344ddfbbead0
Proof reading.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Place any additional packages needed here.  Only include packages
%% which are essential, to avoid problems later. Do NOT use any
%% packages which require e-TeX (for example etoolbox): the e-TeX
%% extensions are not currently available on the ACS conversion
%% servers.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage[version=3]{mhchem} % Formula subscripts using \ce{}
\usepackage[obeyFinal]{easy-todo}
\usepackage{rotating}
\usepackage{epstopdf}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% If issues arise when submitting your manuscript, you may want to
%% un-comment the next line.  This provides information on the
%% version of every file you have used.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%\listfiles

@@

\author{Alexandru Botan}
%\affiliation{The authors are listed in alphaphetical order.}
%\affiliation{The author list is not completed.}
\affiliation[Lyon CNRS]{Institut Lumi\`ere Mati\`ere, UMR5306 Universit\'e Lyon 1-CNRS, Universit\'e de Lyon 69622 Villeurbanne, France}
%
%\author{Andrea Catte}
%\affiliation{University of East Anglia, Norwich, United Kingdom}
\author{Fernando Favela-Rosales}
\affiliation[Mexico]{Departamento de F\'isica, Centro de Investigaci\'on y de Estudios Avanzados del IPN, Apartado Postal 14-740, 07000 M\'exico D.F., M\'exico}
\author{Patrick F. J. Fuchs}
\affiliation[CNRS Paris]{Institut Jacques Monod, CNRS, Universit\'e Paris Diderot, Sorbonne Paris Cit\'e, Paris, France}
%\alsoaffiliation[Diderot University]
%{Institut Jacques Monod, CNRS UMR 7592, Universit{\'e} Paris Diderot, Sorbonne Paris Cit{\'e}, Paris, France}
\author{Matti Javanainen}
\affiliation[Tampere University of Technology]
{Department of Physics, Tampere University of Technology, Tampere, Finland}
\author{Matej Kandu\v{c}}
\affiliation[Freie Universit\"{a}t Berlin] 
{Fachbereich Physik, Freie Universitat Berlin, Berlin, Germany}
\author{Waldemar Kulig}
\affiliation[Tampere University of Technology]
{Department of Physics, Tampere University of Technology, Tampere, Finland}
\author{Antti Lamberg}
\affiliation[Kyoto University]
{Department of Chemical Engineering, Kyoto University, Kyoto, Japan}
\author{Claire Loison}
\affiliation[Lyon CNRS]{Institut Lumi\`ere Mati\`ere, UMR5306 Universit\'e Lyon 1-CNRS, Universit\'e de Lyon 69622 Villeurbanne, France}
\author{Alexander Lyubartsev}
\affiliation[Stockholm University]
{Division of Physical Chemistry, Department of Materials and Environmental Chemistry, Stockholm University, S-106 91 Stockholm, SWEDEN}
\author{Markus S. Miettinen}
\affiliation[Freie Universit\"{a}t Berlin] 
{Fachbereich Physik, Freie Universitat Berlin, Berlin, Germany}
\author{Luca Monticelli}
\affiliation[IBCP] 
{Institut de Biologie et Chimie des Prot\'eines (IBCP), CNRS UMR 5086, Lyon, France}
\author{Jukka M{\"a}{\"a}tt{\"a}}
\affiliation[Aalto University]
{Aalto University, Espoo, Finland}
\author{O. H. Samuli Ollila} 
\email{samuli.ollila@aalto.fi.}
\affiliation[Aalto University]
{Aalto University, Espoo, Finland}
\author{Marius Retegan}
\affiliation[Max Planck]
{Max Planck Institute for Chemical Energy Conversion, Mulheim an der Ruhr, Germany}
\author{Tomasz Rog}
\affiliation[Tampere University of Technology]
{Department of Physics, Tampere University of Technology, Tampere, Finland}
\author{Hubert Santuz}
\affiliation[INSERM]
{INSERM, UMR\_S 1134, DSIMB, Paris, France}
\alsoaffiliation[Diderot]
{Universit\'e Paris Diderot, Sorbonne Paris Cit\'e, UMR\_S 1134, Paris, France}
\alsoaffiliation[INTS]
{Institut National de la Transfusion Sanguine (INTS), Paris, France}
\alsoaffiliation[Labex]
{Laboratoire d'Excellence GR-Ex, Paris, France}
\author{Joona Tynkkynen}
\affiliation[Tampere University of Technology]
{Department of Physics, Tampere University of Technology, Tampere, Finland}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The document title should be given as usual. Some journals require
%% a running title from the author: this should be supplied as an
%% optional argument to \title.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title[An \textsf{achemso} demo]
  {SUPPORTING INFORMATION: Towards atomistic resolution structure of phosphatidylcholine headgroup and glycerol backbone at different ambient conditions\footnote{Publication about results presented in the NMRlipids project}}

@@

%\begin{center}
%{\bf SUPPLEMENTARY INFORMATION}
%\end{center}
\section{Simulation details} 
\subsection{Berger based models}
For the Berger based models we use here the following naming convention: 
Berger - \{{\it molecule name}\} - \{{\it year when model published first time}\} \{{\it citation}\}.
The reason is that there are several different molecular topologies which are using the non-bonded parameters originally
developed by Berger et al.~\cite{berger97}. Thus the common factor in the Berger based models are the non-bonded parameters,
while the molecule specific parameters might somewhat vary. However, the majority of the molecular level topologies are 
relying (especially for the glycerol backbone and headgroup) on the parameters originally introduced by Marrink et al.~\cite{marrink98}.
This is the case for all the Berger based simulations discussed in this work.

POPC simulations at full hydration at 298~K and simulations studying the effect of cholesterol are the same as in previous publications~\cite{ferreira13,ferreira15}.
In these simulation the POPC parameters introduced by Ollila et al~\cite{ollila07a} are used, which are using the non-bonded parameters of Berger~\cite{berger97}
and a molecular topology from Tieleman et al.~\cite{tieleman99} with improved double bond dihedrals by Bachar et al.~\cite{bachar04}. 
Thus they are called Berger-POPC-07~\cite{ollila07a}. The cholesterol model is based on the parameters by H\"oltje et al.~\cite{holtje01} with the
exception that the atom types were changed from CH2/CH3 to LP2/LP3 to avoid overcondensation of the bilayer as suggested in ref.~\cite{tieleman06}.
Since this modification was introduced by Ferreira et al.~\cite{ferreira13}, we call the used cholesterol model as H\"oltje-CHOL-13~\cite{ferreira13}.

For the POPC at 323~K and POPC in low hydration the same force field parameters are used.
For DPPC the implementation of Berger parameters~\cite{berger97} by Peter Tieleman et al. are used~\cite{marrink98}.
For all of these simulations a timestep of 2~fs was used with a leapfrog integrator. Covalent bond lengths were constrained with the LINCS algorithm~\cite{hess97,hess07}. 
Coordinates were written every 10~ps. PME~\cite{darden93,essman95} with real space cut-off of 1.0~nm was used 
for electrostatics. Plain cut-off was used for the Lennard-Jones interactions with a 1.0~nm cut-off.
The neighbor lists with cut-off of 1.0~nm were updated every 5 steps. Temperature was coupled separately
for lipids and water to 298~K using the velocity-rescale method~\cite{bussi07} with coupling constant 0.1~ps.
Pressure was semi-isotropically coupled to the atmospheric pressure with the Berendsen method~\cite{berendsen84}.

@@

{\it DPPC and POPC with 72 lipids.}
%
The starting structures in PDB format were downloaded from the
NIH/NHLBI Laboratory of Computational Biology Membrane Biophysics Section
website (\url{http://www.lobos.nih.gov/mbs/coords.shtml}),
which refers to these as the final structures (for DPPC after 40~ns, POPC after 35~ns)
of the NPT lipid bilayer trajectories used in the original CHARMM36 publication~\cite{klauda10}.
The TIP3P~\cite{jorgensen83} water model was used to solvate the system.
%
The publicly available CHARMM36 parameters in Gromacs format
(September 2013 update: {\tt charmm36\_gmx\_format\_sep13.tgz})
from the MacKerell Lab website
(\url{http://mackerell.umaryland.edu/CHARMM_ff_params.html}) were used.
Timestep of 1~fs was used with the leapfrog integrator. Covalent bonds with hydrogens were constrained with LINCS algorithm~\cite{hess97,hess07}. 
Coordinates were written every 5~ps. PME~\cite{darden93,essman95} with real space cut-off of 1.4~nm was used 
for electrostatics. Lennard-Jones interactions were switched to zero between 0.8~nm and 1.2~nm.
The neighbour lists with a cut-off of 1.4~nm were updated every 5 steps. Temperature was coupled separately
for lipids and water to 303~K using the velocity-rescale method~\cite{bussi07} with coupling constant of 0.2~ps.
Pressure was semi-isotropically coupled to the atmospheric pressure with the Berendsen method~\cite{berendsen84}.

{\it POPC with 128 lipids}
The starting structures for the pure POPC simulations was taken from the Slipids~\cite{jambeck12b} website (\url{http://mmkluster.fos.su.se/slipids/Downloads.html}).
The starting structures for mixed POPC/Cholesterol simulations were constructed with the CHARMM-GUI website~\cite{jo08}. 
They contained 100 POPC/24 cholesterol molecules and 80 POPC/80 cholesterol molecules for
the simulations of 20\% cholesterol and 50\% cholesterol respectively. The TIP3P water model~\cite{jorgensen83} was used to
solvate the system.
The publicly available CHARMM36 forcefield parameters
(\url{http://www.gromacs.org/@api/deki/files/184/=charmm36.ff\\\_4.5.4\_ref.tgz}) 
by Piggot et al. \cite{piggot12} were used. Cholesterol parameters came
from Lim et al. \cite{lim12} and were converted into GROMACS format with the PyTopol tool~\cite{salari15}.  
Single point energy calculation was done to assess the conversion. 
%(At the time I've made the simulations, the charmm36 ff for cholesterol in gromacs format was not available at
%http://mackerell.umaryland.edu/charmm_ff.shtml#gromacs).
Simulations were performed for 200~ns and the last 100~ns was used for the calculations. Timestep of 2~fs was
used with leapfrog integrator. All bond lengths were constrained with LINCS~\cite{hess97,hess07}. Temperature was maintened at
303~K with the velocity-rescale method~\cite{bussi07} and a time constant of 0.2~ps. Pressure was maintained semiisotropically
at 1~bar using the Parrinello--Rahman algorithm~\cite{parrinello81} with a time constant of 1.0~ps. The neighbour list with a cut-off of 1.2~nm
was updated every 10 steps. Lennard-Jones interactions were switched to zero
between 0.8~nm and 1.2~nm. PME~\cite{darden93,essman95} with real space cut-off of 1.2~nm was used for electrostatics.

{\it POPC with 512 lipids.}
%
POPC The starting structures for simulations were constructed with
the CHARMM-GUI website~\cite{jo08}. 512 POPC lipids (256 per leafleat) were 
used for the initial 0\% concentration and they were subsequently 
substituted by cholesterol molecules to fulfil the desired concentration.
The TIP3P~\cite{jorgensen83} water model was used to solvate the system. 
The publicly available port of the CHARMM36 forcefield parameters (\url{mackerell.umaryland.edu/CHARMM_ff_params.html}) 
was used for both POPC and cholesterol. 
Simulations were performed for 170 ns and the last 100 ns was used for the calculations.
Timestep of 1 fs was used with the leapfrog integrator. Covalent bonds with hydrogen 
were constrained with LINCS algorithm. Coordinates were written every 10 ps.
PME with real space cut-off of 1.2 nm was used for electrostatics. Lennard-Jones 
interactions were switched to zero between 0.8 nm and 1.2 nm. The neighbour lists with a
cut-off of 1.2 nm were updated every 5 steps. Temperature was coupled separately for cholesterol, lipids
and water to 298 K using the velocity-rescale method with coupling constant of 0.2 ps.
Pressure was semi-isotropically coupled to the atmospheric pressure with the Berendsen
method.

\subsection{MacRog}
The lipid force field parameters were obtained from the developers and the parameters are presented in Refs.~\citenum{maciejewski14,kulig15b}. 
%correspond to the published DPPC parameters~\cite{maciejewski14} with the inclusion of the double bond parameters~\cite{kulig15}. 
%This inclusion of unsaturated lipid tails will be published in the near future. 
%\todo{You have recent paper where you use POPC/chol simulations with this model: Kulig et al. BBA 1848 (2015) 422-432 http://dx.doi.org/10.1016/j.bbamem.2014.10.032.
%would this be a correct refence here?}
A bilayer with 288 POPC lipids was hydrated with 12600 TIP3P water~\cite{jorgensen83} molecules ($\sim$44/lipid) and simulated for 100~ns with a time step of 2~fs. Data was saved 
every 10~ps and the first 20~ns of the trajectory was discarded from the analysis. 
All bond lengths were constrained with LINCS~\cite{hess97,hess07}. The temperatures of the lipids and the solvent were separately coupled to the Nos\'{e}--Hoover thermostat~\cite{nose84,hoover85} 
with a target temperature of 310~K and a time constant of 0.4~ps. Semi-isotropical pressure coupling to 1~bar was obtained with the Parrinello--Rahman 
barostat~\cite{parrinello81} with a time constant of 1~ps. PME~\cite{darden93,essman95} was employed to calculate the long-range electrostatic interactions. Lennard-Jones interactions were cut off 
at 1~nm and the dispersion correction was applied to both energy and pressure. A neighbour list with a radius of 1~nm was updated every step.

Identical parameters were employed for both full hydration and for the dehydration simulations. The dehydration simulations were also run for 100~ns 
with data saved every 10~ps.

The initial structures for the simulations with 10, 40, 50 and 60 mol\% of cholesterol were obtained by replacing 14, 56, 64 or 72 POPC molecules 
with cholesterol molecules in the initial structure containing 128 POPC molecules. These systems were simulated for 400 ns and the first 200 ns was 
discarded from analysis. Data was saved every 100 ps.

\subsection{GAFFLipid}
The initial structure in Lipidbook \cite{domanski10} had different glycerol backbone isomers in different leaflets. 
To generate the initial structure we took the structure delivered by Slipids developers~\cite{jambeck12b}. Also this structure
had one lipid with different glycerol backbone isomer. This lipid and one lipid from opposite leaflet were removed
after the system was equilibrated.

The force field parameters were generated using files obtained from the Lipidbook website (\url{http://lipidbook.bioch.ox.ac.uk/package/show/id/150.html})~\cite{domanski10}. 
The conversion to GROMACS compatible formats was performed using the acpype tool~\cite{silva12}. The accuracy of the conversion was checked by calculating 
the total energy of a single POPC lipid molecule using the sander program which is part of the AmberTools14 package~\cite{ferrer13} and version 4.6.5 of GROMACS. 
A difference of 0.002 kcal/mol was obtained between the two programs.

Timestep of 2~fs was used in Langevin dynamics with zero friction term and collision frequency of 1.0~ps$^{-1}$. 
Covalent bonds with hydrogens were constrained with the LINCS algorithm~\cite{hess97,hess07}.
Coordinates were written every 10~ps. PME~\cite{darden93,essman95} with a real space cut-off at 1.0~nm was used 
for electrostatics. Plain cut-off with 1~nm was used for Lennard-Jones interactions. 
The neighbour lists with a cut-off of 1.0~nm were updated every 5 steps. 
Pressure was semi-isotropically coupled to a pressure of 1~bar with the Berendsen method~\cite{berendsen84}.

It should be noted that the area per molecule with these settings for the GAFFlipid model was 61.6~\AA$^2$,
while the original publication reported 63.9~\AA$^2$~\cite{dickson12}. However, the same parameters and Amber to Gromacs
conversion reproduced the area per molecule from original publication for the lipid14 model (see next section).

\subsection{Lipid14}
The initial structure was taken directly from the Lipidbook~\cite{domanski10}.
The Amber compatible force field parameters were generated using the tleap program which is integrated in the AmberTools14 package~\cite{ferrer13}. 
A workflow similar to the one used previously for the conversion and validation of the GAFFLipid parameters was followed here. 
As before, a negligible energy difference of 0.003 kcal/mol was obtained between the two programs.

Timestep of 2~fs was used in Langevin dynamics with zero friction term and collision frequency of 1.0~ps$^{-1}$. 
Covalent bonds with hydrogens were constrained with LINCS algorithm~\cite{hess97,hess07}.
Coordinates were written every 10~ps. PME~\cite{darden93,essman95} with real space cut-off of 1.0~nm was used 
for electrostatics. Plain cut-off with 1~nm was used for Lennard-Jones interactions. Dispersion correction
was applied for both energy and pressure. The neighbor lists with a cut-off of 1.0~nm were updated every 5 steps. 
Pressure was semi-isotropically coupled to a pressure of 1~bar with the Berendsen method~\cite{berendsen84}.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Place any additional packages needed here.  Only include packages
%% which are essential, to avoid problems later. Do NOT use any
%% packages which require e-TeX (for example etoolbox): the e-TeX
%% extensions are not currently available on the ACS conversion
%% servers.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\usepackage[version=3]{mhchem} % Formula subscripts using \ce{}
%\usepackage[obeyFinal]{easy-todo}
\usepackage{rotating}
\usepackage{epstopdf}
\usepackage{setspace}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% If issues arise when submitting your manuscript, you may want to
%% un-comment the next line.  This provides information on the
%% version of every file you have used.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%\listfiles

@@

\author{\setstretch{1.05} Alexandru Botan}
%\affiliation{The authors are listed in alphaphetical order.}
%\affiliation{The author list is not completed.}
\affiliation[Lyon CNRS]{\small\setstretch{1.0} Institut Lumi\`ere Mati\`ere, UMR5306 Universit\'e Lyon 1-CNRS, Universit\'e de Lyon 69622 Villeurbanne, France}
%
%\author{Andrea Catte}
%\affiliation{University of East Anglia, Norwich, United Kingdom}
\author{Fernando Favela-Rosales}
\affiliation[Mexico]{Departamento de F\'isica, Centro de Investigaci\'on y de Estudios Avanzados del IPN, Apartado Postal 14-740, 07000 M\'exico D.F., M\'exico}
\author{Patrick F. J. Fuchs}
\affiliation[CNRS Paris]{Institut Jacques Monod, CNRS, Universit\'e Paris Diderot, Sorbonne Paris Cit\'e, Paris, France}
%\alsoaffiliation[Diderot University]
%{Institut Jacques Monod, CNRS UMR 7592, Universit{\'e} Paris Diderot, Sorbonne Paris Cit{\'e}, Paris, France}
\author{Matti Javanainen}
\affiliation[Tampere University of Technology]
{Department of Physics, Tampere University of Technology, Tampere, Finland}
\author{Matej Kandu\v{c}}
\affiliation[Freie Universitat Berlin] 
{Fachbereich Physik, Freie Universit{\"a}t Berlin, Berlin, Germany}
\author{Waldemar Kulig}
\affiliation[Tampere University of Technology]
{Department of Physics, Tampere University of Technology, Tampere, Finland}
\author{Antti Lamberg}
\affiliation[Kyoto University]
{Department of Chemical Engineering, Kyoto University, Kyoto, Japan}
\author{Claire Loison}
\affiliation[Lyon CNRS]{Institut Lumi\`ere Mati\`ere, UMR5306 Universit\'e Lyon 1-CNRS, Universit\'e de Lyon 69622 Villeurbanne, France}
\author{Alexander Lyubartsev}
\affiliation[Stockholm University]
{Division of Physical Chemistry, Department of Materials and Environmental Chemistry, Stockholm University, S-106 91 Stockholm, SWEDEN}
\author{Markus S. Miettinen}
\affiliation[Freie Universitat Berlin] 
{Fachbereich Physik, Freie Universit{\"a}t Berlin, Berlin, Germany}
\author{Luca Monticelli}
\affiliation[IBCP] 
{Institut de Biologie et Chimie des Prot\'eines (IBCP), CNRS UMR 5086, Lyon, France}
\author{Jukka M{\"a}{\"a}tt{\"a}}
\affiliation[Aalto University]
{Aalto University, Espoo, Finland}
\author{O. H. Samuli Ollila} 
\email{samuli.ollila@aalto.fi.}
\affiliation[Aalto University]
{Aalto University, Espoo, Finland}
\author{Marius Retegan}
\affiliation[Max Planck]
{Max Planck Institute for Chemical Energy Conversion, Mulheim an der Ruhr, Germany}
\author{Tomasz Rog}
\affiliation[Tampere University of Technology]
{Department of Physics, Tampere University of Technology, Tampere, Finland}
\author{Hubert Santuz}
\affiliation[INSERM]
{INSERM, UMR\_S 1134, DSIMB, Paris, France}
\alsoaffiliation[Diderot]
{Universit\'e Paris Diderot, Sorbonne Paris Cit\'e, UMR\_S 1134, Paris, France}
\alsoaffiliation[INTS]
{Institut National de la Transfusion Sanguine (INTS), Paris, France}
\alsoaffiliation[Labex]
{Laboratoire d'Excellence GR-Ex, Paris, France}
\author{Joona Tynkkynen}
\affiliation[Tampere University of Technology]
{Department of Physics, Tampere University of Technology, Tampere, Finland}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The document title should be given as usual. Some journals require
%% a running title from the author: this should be supplied as an
%% optional argument to \title.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title[An \textsf{achemso} demo]
  {\setstretch{1.05} SUPPORTING INFORMATION:\\
  Towards atomistic resolution structure of phosphatidylcholine headgroup and glycerol backbone at different ambient conditions\footnote{Publication about results presented in the NMRlipids project (\url{nmrlipids.blogspot.fi})}}

@@

%\begin{center}
%{\bf SUPPLEMENTARY INFORMATION}
%\end{center}
\setstretch{1.025}
\section{Simulation details} 
\subsection{Berger-based models}
For the Berger-based models we used the following naming convention: 
Berger - \{{\it molecule name}\} - \{{\it year when model published first time}\} \{{\it citation}\}.
The reason was that there are several different molecular topologies %which are
using the non-bonded parameters originally
developed by Berger et al.~\cite{berger97}. Thus the common factor in the Berger-based models are the non-bonded parameters,
while the molecule-specific parameters might somewhat vary. However, the majority of the molecular-level topologies  
rely (especially for the glycerol backbone and headgroup) on the parameters originally introduced by Marrink et al.~\cite{marrink98};
this is the case for all the Berger-based simulations discussed in this work.

POPC simulations at full hydration at 298~K, and simulations studying the effect of cholesterol, were the same as in previous publications~\cite{ferreira13,ferreira15}.
In these simulations, the POPC parameters introduced in 2007 by Ollila et al.~\cite{ollila07a} were used (the non-bonded parameters of Berger et al.~\cite{berger97}
and a molecular topology from Tieleman et al.~\cite{tieleman99} with improved double bond dihedrals by Bachar et al.~\cite{bachar04}), thus
they were called Berger-POPC-07~\cite{ollila07a}. The cholesterol model was based on the parameters by H\"oltje et al.~\cite{holtje01} with the
exception that the atom types were changed from CH2/CH3 to LP2/LP3 to avoid overcondensation of the bilayer as suggested in Ref.~\citenum{tieleman06}.
Since this modification was introduced in 2013 by Ferreira et al.~\cite{ferreira13}, we call this cholesterol model H\"oltje-CHOL-13~\cite{ferreira13}.

For the POPC at 323~K and POPC at low hydration, the same force field parameters were used.
For DPPC, the implementation of Berger parameters~\cite{berger97} by Peter Tieleman et al. was used~\cite{marrink98}.
For all  these simulations, a timestep of 2~fs was used with a leapfrog integrator. Covalent bond lengths were constrained with the LINCS algorithm~\cite{hess97,hess07}. 
Coordinates were written every 10~ps. PME~\cite{darden93,essman95} with a real space cut-off of 1.0~nm was used 
for electrostatics. Plain cut-off was used for the Lennard-Jones interactions, with a 1.0~nm cut-off.
The neighbor lists with a cut-off of 1.0~nm were updated every 5 steps. Temperature was coupled separately
for lipids and water to 298~K using the velocity-rescale method~\cite{bussi07} with coupling constant 0.1~ps.
Pressure was semi-isotropically coupled to the atmospheric pressure with the Berendsen method~\cite{berendsen84}.

@@

{\it DPPC and POPC with 72 lipids.}
%
The starting structures in PDB format were downloaded from the
NIH/NHLBI Laboratory of Computational Biology Membrane Biophysics Section
website (\url{http://www.lobos.nih.gov/mbs/coords.shtml}),
which refers to these as the final structures (for DPPC after 40~ns, POPC after 35~ns)
of the NPT lipid bilayer trajectories used in the original CHARMM36 publication~\cite{klauda10}.
The TIP3P~\cite{jorgensen83} water model was used to solvate the system.
%
The publicly available CHARMM36 parameters in Gromacs format
(September 2013 update: {\tt charmm36\_gmx\_format\_sep13.tgz})
from the MacKerell Lab website
(\url{http://mackerell.umaryland.edu/CHARMM_ff_params.html}) were used.
Timestep of 1~fs was used with the leapfrog integrator. Covalent bonds with hydrogens were constrained with the LINCS algorithm~\cite{hess97,hess07}. 
Coordinates were written every 5~ps. PME~\cite{darden93,essman95} with real space cut-off of 1.4~nm was used 
for electrostatics. Lennard-Jones interactions were switched to zero between 0.8~nm and 1.2~nm.
The neighbor lists with a cut-off of 1.4~nm were updated every 5 steps. Temperature was coupled separately
for lipids and water to 303~K using the velocity-rescale method~\cite{bussi07} with a coupling constant of 0.2~ps.
Pressure was semi-isotropically coupled to the atmospheric pressure with the Berendsen method~\cite{berendsen84}.

{\it POPC with 128 lipids.}
The starting structures for the pure POPC simulation was taken from the Slipids~\cite{jambeck12b} website (\url{http://mmkluster.fos.su.se/slipids/Downloads.html}).
The starting structures for the mixed POPC/Cholesterol simulations were constructed with the CHARMM-GUI website~\cite{jo08}. 
They contained 100 POPC~/~24 cholesterol molecules and 80 POPC~/~80 cholesterol molecules for
the simulations of 20\% cholesterol and 50\% cholesterol, respectively. The TIP3P water model~\cite{jorgensen83} was used to
solvate the system.
The publicly available CHARMM36 forcefield parameters
(\url{http://www.gromacs.org/@api/deki/files/184/=charmm36.ff\\\_4.5.4\_ref.tgz}) 
by Piggot et al. \cite{piggot12} were used. Cholesterol parameters came
from Lim et al. \cite{lim12}, and were converted into GROMACS format with the PyTopol tool~\cite{salari15}.  
Single-point energy calculation was done to assess the conversion. 
%(At the time I've made the simulations, the charmm36 ff for cholesterol in gromacs format was not available at
%http://mackerell.umaryland.edu/charmm_ff.shtml#gromacs).
Simulations were performed for 200~ns and the last 100~ns was used for the analysis. Timestep of 2~fs was
used with leapfrog integrator. All bond lengths were constrained with LINCS~\cite{hess97,hess07}. Temperature was maintened at
303~K using the velocity-rescale method~\cite{bussi07} with a time constant of 0.2~ps. Pressure was maintained semi-isotropically
at 1~bar using the Parrinello--Rahman algorithm~\cite{parrinello81} with a time constant of 1.0~ps. The neighbor list with a cut-off of 1.2~nm
was updated every 10 steps. Lennard-Jones interactions were switched to zero
between 0.8~nm and 1.2~nm. PME~\cite{darden93,essman95} with a real space cut-off of 1.2~nm was used for electrostatics.

{\it POPC with 512 lipids.}
%
The starting structures for simulations were constructed with
the CHARMM-GUI website~\cite{jo08}. 512 POPC lipids (256 per leafleat) were 
used for the initial 0\% cholesterol content, and they were subsequently 
substituted by cholesterol molecules to obtain the desired content.
The TIP3P~\cite{jorgensen83} water model was used to solvate the system. 
The publicly available port of the CHARMM36~\cite{klauda10} forcefield parameters (\url{mackerell.umaryland.edu/CHARMM_ff_params.html}) 
was used for both the POPC and cholesterol. 
Simulations were performed for 170~ns, and the last 100~ns was used for the analysis.
Timestep of 1~fs was used with the leapfrog integrator. Covalent bonds with hydrogen 
were constrained with the LINCS~\cite{hess97,hess07} algorithm. Coordinates were written every 10~ps.
PME~\cite{darden93,essman95} with real space cut-off of 1.2~nm was used for electrostatics. Lennard-Jones 
interactions were switched to zero between 0.8~nm and 1.2~nm. The neighbor lists with a
cut-off of 1.2~nm were updated every 5 steps. Temperature was coupled separately for cholesterol, lipids
and water to 298~K using the velocity-rescale method~\cite{bussi07} with coupling constant of 0.2~ps.
Pressure was semi-isotropically coupled to the atmospheric pressure with the Berendsen
method~\cite{berendsen84}.

\subsection{MacRog}
The lipid force field parameters were obtained from the developers; the parameters are presented in Refs.~\citenum{maciejewski14,kulig15b}. 
%correspond to the published DPPC parameters~\cite{maciejewski14} with the inclusion of the double bond parameters~\cite{kulig15}. 
%This inclusion of unsaturated lipid tails will be published in the near future. 
%\todo{You have recent paper where you use POPC/chol simulations with this model: Kulig et al. BBA 1848 (2015) 422-432 http://dx.doi.org/10.1016/j.bbamem.2014.10.032.
%would this be a correct refence here?}
A bilayer with 288 POPC lipids was hydrated with 12600 TIP3P water~\cite{jorgensen83} molecules ($\sim$44/lipid) and simulated for 100~ns with a time step of 2~fs. Data was saved 
every 10~ps and the first 20~ns of the trajectory was discarded from the analysis. 
All bond lengths were constrained with LINCS~\cite{hess97,hess07}. The temperatures of the lipids and the solvent were separately coupled to the Nos\'{e}--Hoover thermostat~\cite{nose84,hoover85} 
with a target temperature of 310~K and a time constant of 0.4~ps. Semi-isotropical pressure coupling to 1~bar was obtained with the Parrinello--Rahman 
barostat~\cite{parrinello81} with a time constant of 1~ps. PME~\cite{darden93,essman95} was employed to calculate the long-range electrostatic interactions. Lennard-Jones interactions were cut off 
at 1~nm and the dispersion correction was applied to both energy and pressure. A neighbour list with a radius of 1~nm was updated every step.

Identical parameters were employed for the full hydration and the dehydration simulations. The dehydration simulations were also run for 100~ns 
with data saved every 10~ps.

The initial structures for the simulations with 10, 40, 50 and 60 mol-\% of cholesterol were obtained by replacing 14, 56, 64 or 72 POPC molecules 
with cholesterol molecules in the initial structure containing 128 POPC molecules. These systems were simulated for 400 ns and the first 200 ns was 
discarded from analysis. Data was saved every 100 ps.

\subsection{GAFFlipid}
The initial structure in Lipidbook \cite{domanski10} had different glycerol backbone isomers in different leaflets. 
To generate the initial structure, we took the structure delivered by the Slipids developers~\cite{jambeck12b}. Also this structure
had one lipid with different glycerol backbone isomer. This lipid and one lipid from opposite leaflet were removed
after the system was equilibrated.

The force field parameters were generated using files obtained from the Lipidbook website (\url{http://lipidbook.bioch.ox.ac.uk/package/show/id/150.html})~\cite{domanski10}. 
The conversion to GROMACS compatible formats was performed using the acpype tool~\cite{silva12}. Accuracy of the conversion was checked by calculating 
the total energy of a single POPC lipid molecule using the sander program, which is part of the AmberTools14 package~\cite{ferrer13}, and version 4.6.5 of GROMACS. 
A difference of 0.002 kcal/mol was obtained between the two programs.

Timestep of 2~fs was used in Langevin dynamics with zero friction term and collision frequency of 1.0~ps$^{-1}$. 
Covalent bonds with hydrogens were constrained with the LINCS algorithm~\cite{hess97,hess07}.
Coordinates were written every 10~ps. PME~\cite{darden93,essman95} with a real space cut-off at 1.0~nm was used 
for electrostatics. Plain cut-off at 1~nm was used for Lennard-Jones interactions. 
The neighbor lists with a cut-off of 1.0~nm were updated every 5 steps. 
Pressure was semi-isotropically coupled to 1~bar with the Berendsen method~\cite{berendsen84}.

It should be noted that the area per molecule with these settings for the GAFFlipid model was 61.6~\AA$^2$,
while the original publication~\cite{dickson12} reported 63.9~\AA$^2$. Notably, however, the same parameters and Amber-to-Gromacs
conversion procedure reproduced the area per molecule from the original publication~\cite{dickson14} for the Lipid14 model (see the Subsec.~Lipid14).

\subsection{Lipid14}
The initial structure was taken directly from the Lipidbook~\cite{domanski10}.
The Amber-compatible force field parameters were generated using the tleap program, which is integrated in the AmberTools14 package~\cite{ferrer13}. 
A workflow similar to the one used for the conversion and validation of the GAFFlipid parameters was followed;
here a negligible energy difference of 0.003 kcal/mol was obtained between sander and GROMACS.

Timestep of 2~fs was used in Langevin dynamics with zero friction term and a collision frequency of 1.0~ps$^{-1}$. 
Covalent bonds with hydrogens were constrained with the LINCS algorithm~\cite{hess97,hess07}.
Coordinates were written every 10~ps. PME~\cite{darden93,essman95} with real space cut-off at 1.0~nm was used 
for electrostatics. Plain cut-off with 1~nm was used for Lennard-Jones interactions. Dispersion correction
was applied for both energy and pressure. The neighbor lists with a cut-off of 1.0~nm were updated every 5 steps. 
Pressure was semi-isotropically coupled to a pressure of 1~bar with the Berendsen method~\cite{berendsen84}.
